Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-05-05 paper-correlations.tex
Fix a typo
Update execution plots

For surface-surface and surface-void functions we use an edge detection filter
and then calculate either autocorrelation of the edge or cross-correlation of
the edge and the void phase. The edge is extracted by convolving an input with a
short low-pass filter $H$. It has a width $7$ and all its coefficients with
exception of the central coefficient are inversely proportional to distance
from the center:
\begin{equation}
  \begin{aligned}
    H_{ij} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2}} & \quad i \ne 3, j \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right. \\
    H_{ijk} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2 + (k-3)^3}} & \quad i \ne 3, j \ne 3, k
      \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right.
  \end{aligned}
  \label{eq:filter-7x7}
\end{equation}
where $C$ is such that all coefficients of $H'$ sum to zero, and $S$ equals to
$30.45849$ in 2D and $172.96232$ in 3D case.

@@

\section{Execution times of our algorithms}
\label{sec:efficiency}
To measure execution times of our algorithms we use a machine with the following
configuration:\textcolor{red}{Must be configuration of Asus notebook}
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 5 1600X
\item \textbf{Memory}: 32~Gb DDR4-2400 memory
\item \textbf{GPU}: Nvidia Tesla V100 with 32~Gb video memory
\end{itemize}

We generate two (resp. three) dimensional square (resp. cubic) binary arrays
with different sides. An array used for measurements contains two solid disks
(resp. balls) surrounded by void phase, therefore it is excected for $C_2$
function to perform two times slower than $S_2$.

To measure efficiency of \code{Directional} module we generate ten arrays with
a side of a square $s = 1000, 2000, \dots, 10000$ pixels for two-dimensional
case and ten arrays with a side of a cube $s = 50, 100, \dots, 500$ voxels for
three-dimensional case and calculate execution time for each function. All
functions are calculated along axial directions using periodic boundary
conditions.

With the \code{Map} module we follow the same procedure. Many functions in
\code{Map} can utilize Nvidia's GPUs using the CUDA library. For example,
two-point correlation function uses FFT transform optimized for Nvidia GPUs and
provided in \code{CUDA.FFT} package. We provide execution times for both CPU and
GPU. The choice between CPU or GPU version is made depending on type of input
array. If it's \code{CuArray} rather than generic \code{AbstractArray} then CUDA
implementation is chosen. \textcolor{red}{Please note, that CPU versions are
  much slower, so we choose a different sides of squares
  ($s = 700, 1350, \dots, 7000$) and cubes ($s = 10, 45, \dots, 350$)
  for evaluation of the performance.}

The measurements of execution time is on \cref{fig:timings}. As you can see, GPU
versions (when available) are faster than CPU versions and limited only by
amount of video memory. Functions in \code{Directional} are the slowest among
other implementations, but require less memory.

>>>>

For surface-surface and surface-void functions we use an edge detection filter
and then calculate either autocorrelation of the edge or cross-correlation of
the edge and the void phase. The edge is extracted by convolving an input with a
short high-pass filter $H$. It has a width $7$ and all its coefficients with
exception of the central coefficient are inversely proportional to distance
from the center:
\begin{equation}
  \begin{aligned}
    H_{ij} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2}} & \quad i \ne 3, j \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right. \\
    H_{ijk} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2 + (k-3)^3}} & \quad i \ne 3, j \ne 3, k
      \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right.
  \end{aligned}
  \label{eq:filter-7x7}
\end{equation}
where $C$ is such that all coefficients of $H'$ sum to zero, and $S$ equals to
$30.45849$ in 2D and $172.96232$ in 3D case.

@@

\section{Execution times of our algorithms}
\label{sec:efficiency}
To measure execution times of our algorithms we use a machine with the following
configuration:
\begin{itemize}
\item \textbf{CPU}: AMD Ryzen 7 5800X
\item \textbf{Memory}: 32~Gb DDR4 memory
\item \textbf{GPU}: Nvidia Tesla V100 with 32~Gb video memory
\end{itemize}

We generate two (resp. three) dimensional square (resp. cubic) binary arrays
with different sides. An array used for measurements contains two solid disks
(resp. balls) surrounded by void phase.

To measure efficiency of our code we generate six arrays with a side of a square
$s = 1000, 2000, \dots, 6000$ pixels for two-dimensional case and six arrays
with a side of a cube $s = 50, 100, \dots, 300$ voxels for three-dimensional
case and calculate execution time for each function. All functions are
calculated using periodic boundary conditions. Functions which use per-slice
directional approach are calculated along axial directions.

Functions which calculate full correlation maps are suitable for running on CPU
and GPU, so we provide execution times for both. The cluster function still uses
an algorithm for closest component labeling which works only on CPU, so there is
a potentional for improvement for this function.

The measurements of execution time is on \cref{fig:timings}. As you can see, GPU
versions (when available) are faster than CPU versions and limited only by
amount of video memory. Functions in module \code{Directional} (which utilize
per-slice directional approach) are the slowest among other implementations, but
require less memory.
